---
Title: "Studying the relationship betwen climate change and children's wellbeing"
Author: Hernando Grueso
Output: climate_children
---

Creation: 09/05/2023
Update: 25/08/2023
Data Sources: 
1) Flood maps from Nigeria: https://global-flood-database.cloudtostreet.ai/
2) DHS:
3) VACS:
4) MICS:

A. Install Required Packages
```{r}
# Install Required Packages
if(!requireNamespace("raster")) install.packages( "raster" )
if(!requireNamespace("MASS")) install.packages( "MASS" )
if(!requireNamespace("rgdal")) install.packages( "rgdal" )
if(!requireNamespace("spatstat")) install.packages( "spatstat" )
if(!requireNamespace("sp")) install.packages( "sp" )
if(!requireNamespace("rnaturalearth")) install.packages( "rnaturalearth" )
if(!requireNamespace("rnaturalearthdata")) install.packages( "rnaturalearthdata" )
if(!requireNamespace("dplyr")) install.packages( "dplyr" )
if(!requireNamespace("sf")) install.packages( "sf" )
if(!requireNamespace("RColorBrewer")) install.packages( "RColorBrewer" )
if(!requireNamespace("leaflet")) install.packages( "leaflet" )
if(!requireNamespace("htmlwidgets")) install.packages( "htmlwidgets" )
if(!requireNamespace("ggplot2")) install.packages( "ggplot2" )
if(!requireNamespace("stargazer")) install.packages( "stargazer" )
if(!requireNamespace("Matching")) install.packages( "Matching" )
if(!requireNamespace("rgenoud")) install.packages( "rgenoud" )
if(!requireNamespace("writexl")) install.packages( "writexl" )
if(!requireNamespace("rnaturalearth")) install.packages( "rnaturalearth" )
if(!requireNamespace("glm2")) install.packages( "glm2" )

```

B. Define Directory
```{r}
# Defining Directory:
groupDir <- "C:/Users/hgruesohurtado/Desktop/Climate Change"

```

C. Import Data
```{r}
# Import GeoTiff file
library(raster)
raster_brick <- brick("Floods/2018/DFO_4667_From_20180820_to_20181002.tif")
print(raster_brick)

# Convert to a dataframe
df <- as.data.frame(raster_brick, xy = TRUE)
head(df)

# Import DHS geo-codes
library(readr)
dhs <- read_csv("Save the Children/Child poverty and climate/dhsgps_combined_sep22.csv", show_col_types = FALSE)
dhs <- subset(dhs, dhs$DHSYEAR==2018)

# Import DHS data 2018
dhs_df <- read_csv("Save the Children/Child poverty and climate/Microdata_DHS_Apr2023.csv", show_col_types = FALSE)
dhs_df <- subset(dhs_df, dhs_df$year_interview==2018)
```

D. Merge DHS MPI & GEO-Codes
```{r}
# Only keep Nigeria in DHS
dhs <- subset(dhs, dhs$DHSCC=="NG")
dhs_df <- subset(dhs_df, dhs_df$countrycode=="NGA")

# Merge at the cluster level (geo-codes are defined at the cluster level)
dhs_df <- merge(dhs_df, dhs, by = "cluster")

# Delete data frames that will not be used
rm(df)
```

E.1 Plot Nigeria flooding map
```{r}
# Plot the raster map with custom settings
plot(raster_brick[[1]],
     main = "Flooded Areas: Nigeria, 2018",
     legend = TRUE,
     legend.args = list(text = "Flooded"))
```

E.2 Plot Nigeria flooding map: Durantion
```{r}
# Plot the raster map with custom settings
plot(raster_brick[[2]],
     main = "Flood Duration Map: Nigeria, 2018",
     legend = TRUE,
     legend.args = list(text = "Duration"))
```

E.3 Plot Nigeria flooding map: Durantion
```{r}
# Plot the raster map with custom settings
plot(raster_brick[[5]],
     main = "Permanent Water Map: Nigeria, 2018",
     legend = TRUE,
     legend.args = list(text = "P. Water"))
```

F.1 Plot Severe MPI=1 DHS households on flooding map  
```{r}
# Plot the raster layer
plot(raster_brick[[1]],
     main = "Flood Map: Nigeria, 2018 (Child Deprived: Severe)",
     legend = TRUE,
     legend.args = list(text = "Flooded"))

# Create subset of severe poor households
dhs_poor_sev <- subset(dhs_df, dhs_df$deprived_sev==1)
  
# Add DHS data points to the plot
points(dhs_poor_sev$LONGNUM, dhs_poor_sev$LATNUM, col = "blue", pch = 20)
```

F.2 Plot Moderate MPI=0 DHS households on flooding map  
```{r}
# Plot the raster layer
plot(raster_brick[[1]],
     main = "Flood Map: Nigeria, 2018 (Child Non-Deprived: Moderate)",
     legend = TRUE,
     legend.args = list(text = "Flooded"))

# Create subset of moderately poor households
dhs_poor_mod <- subset(dhs_df, dhs_df$deprived_mod==0)
  
# Add DHS data points to the plot
points(dhs_poor_mod$LONGNUM, dhs_poor_mod$LATNUM, col = "purple", pch = 20)

```

G.1 Plot Severe MPI=0 DHS households on flooding map  
```{r}
# Plot the raster layer
plot(raster_brick[[1]],
     main = "Flood Map: Nigeria, 2018 (Child Non-Deprived: Severe)",
     legend = TRUE,
     legend.args = list(text = "Flooded"))

# Create subset of severe poor households
dhs_poor_sev <- subset(dhs_df, dhs_df$deprived_sev==0)
  
# Add DHS data points to the plot
points(dhs_poor_sev$LONGNUM, dhs_poor_sev$LATNUM, col = "blue", pch = 20)

```

G.2 Plot Flooding Map & DHS Households Deprived on: Housing (moderate)  
```{r}
# Plot the raster layer
plot(raster_brick[[1]],
     main = "Flood Map: Nigeria, 2018 (Child Non-Deprived: Moderate)",
     legend = TRUE,
     legend.args = list(text = "Flooded"))

# Create subset of moderately poor households
dhs_poor_mod <- subset(dhs_df, dhs_df$dep_housing_mod==1)
  
# Add DHS data points to the plot
points(dhs_poor_mod$LONGNUM, dhs_poor_mod$LATNUM, col = "orange", pch = 20)

```

H.1 Plot Flooding Map & DHS Households Deprived on: Water
```{r}
# Plot the raster layer
plot(raster_brick[[1]],
     main = "Flood Map: Nigeria, 2018 (Water Deprivation: Moderate)",
     legend = TRUE,
     legend.args = list(text = "Flooded"))

# Create subset of moderately poor households
dhs_poor_mod <- subset(dhs_df, dhs_df$dep_water_mod==1)
  
# Add DHS data points to the plot
points(dhs_poor_mod$LONGNUM, dhs_poor_mod$LATNUM, col = "blue", pch = 20)

```

H.2 Plot Flooding Map & DHS Households Deprived on: Sanitation
```{r}
# Plot the raster layer
plot(raster_brick[[1]],
     main = "Flood Map: Nigeria, 2018 (Sanitation Deprivation: Moderate)",
     legend = TRUE,
     legend.args = list(text = "Flooded"))

# Create subset of moderately poor households
dhs_poor_mod <- subset(dhs_df, dhs_df$dep_sanitation_mod==1)
  
# Add DHS data points to the plot
points(dhs_poor_mod$LONGNUM, dhs_poor_mod$LATNUM, col = "red", pch = 20)

```

H.3 Plot Flooding Map & DHS Households Deprived on: Nutrition
```{r}
# Plot the raster layer
plot(raster_brick[[1]],
     main = "Flood Map: Nigeria, 2018 (Nutrition Deprivation: Moderate)",
     legend = TRUE,
     legend.args = list(text = "Flooded"))

# Create subset of moderately poor households
dhs_poor_mod <- subset(dhs_df, dhs_df$dep_nutrition_mod==1)
  
# Add DHS data points to the plot
points(dhs_poor_mod$LONGNUM, dhs_poor_mod$LATNUM, col = "purple", pch = 20)

```

H.4 Plot Flooding Map & DHS Households Deprived on: Health
```{r}
# Plot the raster layer
plot(raster_brick[[1]],
     main = "Flood Map: Nigeria, 2018 (Health Deprivation: Moderate)",
     legend = TRUE,
     legend.args = list(text = "Flooded"))

# Create subset of moderately poor households
dhs_poor_mod <- subset(dhs_df, dhs_df$dep_health_mod==1)
  
# Add DHS data points to the plot
points(dhs_poor_mod$LONGNUM, dhs_poor_mod$LATNUM, col = "orange", pch = 20)

```

H.6 Plot Flooding Map & DHS Households Deprived on: Education
```{r}
# Plot the raster layer
plot(raster_brick[[1]],
     main = "Flood Map: Nigeria, 2018 (Education Deprivation: Moderate)",
     legend = TRUE,
     legend.args = list(text = "Flooded"))

# Create subset of moderately poor households
dhs_poor_mod <- subset(dhs_df, dhs_df$dep_education_mod==1)
  
# Add DHS data points to the plot
points(dhs_poor_mod$LONGNUM, dhs_poor_mod$LATNUM, col = "yellow", pch = 20)

```

I.1A Map of Poverty Severity at the Cluster Level
```{r}
library(raster)
library(ggplot2)
library(scales)
library(RColorBrewer)

# Define desired extent values (adjust as needed)
min_longitude <- 0
max_longitude <- 17
min_latitude <- 0
max_latitude <- 17

# Plot the raster data with a lighter color palette
color_blend_func <- colorRampPalette(colors = c("green", brewer.pal(9, "RdYlBu")))
plot(raster_brick[[1]], col = color_blend_func(256), axes = TRUE, main = 'Flood Map: Nigeria, 2018 with Cluster Poverty Severity', xlim=c(min_longitude, max_longitude), ylim=c(min_latitude, max_latitude), asp=1)

# Create a color palette for the points
color_func <- colorRampPalette(colors = c("red", "red"))
color_indices <- round(dhs_df$deprived_sev * 100)
color_indices[color_indices > 100] <- 100
color_indices[color_indices < 1] <- 1
colors <- alpha(color_func(100)[color_indices], 0.7)

# Add the clusters' locations to the plot with small size and red color
points(dhs_df$LONGNUM, dhs_df$LATNUM, col = colors, pch = 20, cex = 0.5)

```

I.1B Map of Poverty Severity at the Cluster Level
```{r}
# Required libraries
library(raster)
library(ggplot2)
library(scales)
library(RColorBrewer)
library(rnaturalearth)

# Load the Nigeria shapefile
nigeria <- ne_countries(scale = "medium", returnclass = "sf", country = "Nigeria")

# Color palette for the raster data
color_blend_func <- colorRampPalette(colors = c("green", brewer.pal(9, "RdYlBu")))

# Create a color palette for the points
color_func <- colorRampPalette(colors = c("red", "red"))
color_indices <- round(dhs_df$deprived_sev * 100)
color_indices[color_indices > 100] <- 100
color_indices[color_indices < 1] <- 1
colors <- alpha(color_func(100)[color_indices], 0.7)

# Create the plot using ggplot2
plot <- ggplot() +
  geom_raster(data = as.data.frame(raster_brick[[1]], xy = TRUE), aes(x = x, y = y, fill = flooded), interpolate = TRUE) +
  scale_fill_gradientn(colors = color_blend_func(256), na.value = "white") +
  geom_sf(data = nigeria, fill = NA, color = "black") +  # Bring back Nigeria's border with black color
  geom_point(data = dhs_df, aes(x = LONGNUM, y = LATNUM, color = deprived_sev), size = 0.5, alpha = 0.7) +
  scale_color_gradientn(colors = color_func(100)) +
  coord_sf(ylim = c(4, 16.5), expand = FALSE) +
  labs(
    title = 'Flood Map: Nigeria, 2018 with Cluster Poverty Severity',
    x = "Longitude",
    y = "Latitude"
  ) +
  theme_minimal() +
  theme(
    panel.background = element_blank(),
    plot.background = element_blank(),
    panel.grid = element_blank(),
    panel.border = element_blank(),  # Remove default box around the plot area
    axis.line = element_blank()  # Remove axis lines
  ) +
  guides(color = FALSE)

print(plot)

```

I.2 Geo-Spatial Distribution of Poverty
```{r}
library(tidyverse)

# Collapse poverty insidence at the cluster level
dhs_cluster <- dhs_df %>%
  filter(LATNUM != 0 & LONGNUM != 0) %>%
  group_by(cluster) %>%
  summarize(avg_deprived_sev = mean(deprived_sev),
            LATNUM = first(LATNUM),
            LONGNUM = first(LONGNUM))

# Plot the data
library(ggplot2)
ggplot(dhs_cluster, aes(x = LONGNUM, y = LATNUM)) +
  geom_point(aes(color = avg_deprived_sev)) +
  scale_color_gradient(low = "yellow", high = "red") +
  labs(title = "Geo-Spatial Distribution of Poverty",
       x = "Longitude", 
       y = "Latitude",
       color = "Average Deprivation Level") +
  theme_minimal()
```

I.3 Geo-Spatial Distribution of Wealth
```{r}
# Collapse poverty incidence at the cluster level
dhs_cluster <- dhs_df %>%
  filter(LATNUM != 0 & LONGNUM != 0) %>%
  group_by(cluster) %>%
  summarize(wealth = mean(wealth),
            LATNUM = first(LATNUM),
            LONGNUM = first(LONGNUM))

# Load the necessary library
library(ggplot2)

# Check the data
head(dhs_cluster)

# Plot the data
ggplot(dhs_cluster, aes(x = LONGNUM, y = LATNUM)) +
  geom_point(aes(color = wealth)) +
  scale_color_gradient(low = "red", high = "yellow") +
  labs(title = "Geo-Spatial Distribution of Wealth",
       x = "Longitude", 
       y = "Latitude",
       color = "Average Wealth Level") +
  theme_minimal()
```

J.1 Kernel Density Estimate
```{r}
# Convert the raster to points
points <- rasterToPoints(raster_brick[[1]])

# Perform kernel density estimation
kde <- density(points)

# Plot the kernel density estimate
plot(kde)

```

J.2 Calculate KDE 1x1 and 11x11
```{r}
library(raster)
library(RColorBrewer)
library(sp)
library(lattice)
library(grid)

# Convert the raster to a binary mask
flooded_mask <- raster_brick[[1]] > 0

# Calculate density with 1x1 bandwidth
bandwidth1 <- matrix(3, nrow = 3, ncol = 3)
density1 <- focal(flooded_mask, w = bandwidth1, fun = sum)

# Calculate density with 17x17 bandwidth
bandwidth11 <- matrix(3, nrow = 17, ncol = 17)
density11 <- focal(flooded_mask, w = bandwidth11, fun = sum)

# Identify the raster with fewer non-NA values
if (sum(!is.na(density1[])) < sum(!is.na(density11[]))) {
  mask_raster <- density1
} else {
  mask_raster <- density11
}

# Use the mask_raster to mask both density rasters
density1_aligned <- mask(density1, mask_raster)
density11_aligned <- mask(density11, mask_raster)

# Convert the rasters to points (data frames)
density1_points <- rasterToPoints(density1_aligned)
density11_points <- rasterToPoints(density11_aligned)
duration_points <- rasterToPoints(raster_brick[[2]])
permanent_water_points <- rasterToPoints(raster_brick[[5]])

# Mask all the rasters by the mask_raster
density1_masked <- mask(density1_aligned, mask_raster)
density11_masked <- mask(density11_aligned, mask_raster)
duration_masked <- mask(raster_brick[[2]], mask_raster)
permanent_water_masked <- mask(raster_brick[[5]], mask_raster)

# Convert the masked rasters to points
density1_points <- rasterToPoints(density1_masked)
density11_points <- rasterToPoints(density11_masked)
duration_points <- rasterToPoints(duration_masked)
permanent_water_points <- rasterToPoints(permanent_water_masked)

# Now, combine into one data frame
df_flood18 <- data.frame(
  Longitude = density1_points[,1],
  Latitude = density1_points[,2],
  Density1 = density1_points[,3],
  Density11 = density11_points[,3],
  Duration = duration_points[,3],
  PermanentWater = permanent_water_points[,3]
)

# Combine the two density rasters into a multilayer raster
density_stack <- stack(density1_aligned, density11_aligned)

# Create a custom multi-color palette
my_palette <- colorRampPalette(c(brewer.pal(3, "Blues"), brewer.pal(3, "Greens"), brewer.pal(3, "Reds")))

# Create a custom plot function using levelplot (from the lattice package)
# Plot the two density rasters with the custom color palette
spplot_obj <- spplot(density_stack, names.attr = c("Bandwidth 3x3", "Bandwidth 17x17"), col.regions = my_palette(100))

# Print the spplot object
print(spplot_obj)

# Add a title
grid.text("Kernel Density Estimates: Flooding Nigeria, 2018", x = 0.5, y = 0.98, gp = gpar(fontsize = 15))

```

J.3 Merge KDE and DHS
```{r}
# Rename variable before merging
names(df_flood18)[names(df_flood18) == "Longitude"] <- "LONGNUM"
names(df_flood18)[names(df_flood18) == "Latitude"] <- "LATNUM"
names(df_flood18)[names(df_flood18) == "Density1"] <- "flood_kde_1"
names(df_flood18)[names(df_flood18) == "Density11"] <- "flood_kde_11"

# Merge at the cluster level (geo-codes are defined at the cluster level)
dhs_flood <- merge(dhs_df, df_flood18, by = c("LATNUM", "LONGNUM"))

# Convert datasets to sf objects
library(sf)
dataset1_sf <- st_as_sf(dhs_df, coords = c("LONGNUM", "LATNUM"))
dataset2_sf <- st_as_sf(df_flood18, coords = c("LONGNUM", "LATNUM"))

# Perform a spatial join based on proximity, only include matched observations
dhs_flood <- st_join(dataset1_sf, dataset2_sf, join = st_nearest_feature, left = FALSE, dist = 100)  # Adjust the distance threshold as needed

# Check the number of approximate matches
num_approx_matches <- nrow(dhs_flood)
print(dhs_flood)
```

J.4A Descriptive Statistics
```{r}
library(stargazer)
library(dplyr)

# Create the new variable "rural"
dhs_flood$rural <- ifelse(dhs_flood$URBAN_RURA == "R", 1, 0)

# Create the new variable "female"
dhs_flood$female <- ifelse(dhs_flood$sex == "Female", 1, 0)

# Create a subset for only hosehold located in flooded areas
dhs_flood <- subset(dhs_flood, flood_kde_11 != 0)

# Subset
flood_yes <- dhs_flood[dhs_flood$flood_kde_1>0, ]
flood_no <- dhs_flood[dhs_flood$flood_kde_1==0, ]

flood_yes <- flood_yes[c("deprived_sev", "deprived_mod", "PermanentWater", "Duration", "age", "rural", "female", "wealth")]
flood_no <- flood_no[c("deprived_sev", "deprived_mod", "PermanentWater", "Duration", "age", "rural", "female", "wealth")]

flood_yes$geometry <- NULL
flood_no$geometry <- NULL

# Compute descriptive stats
mean_df_r <- sapply(flood_yes, mean)
sd_df_r <- sapply(flood_yes, sd)
obs_df_r <- sapply(flood_yes, function(x) sum(!is.na(x)))

mean_df_u <- sapply(flood_no, mean)
sd_df_u <- sapply(flood_no, sd)
obs_df_u <- sapply(flood_no, function(x) sum(!is.na(x)))

stargazer(as.data.frame(flood_yes), 
          type = "text", 
          title="Flooding Area (KDE, 1X1): Yes", 
          digits=2, 
          out="desc_stats_flood_yes.txt")

stargazer(as.data.frame(flood_no), 
          type = "text", 
          title="Flooding Area (KDE, 1X1): No", 
          digits=2, 
          out="desc_stats_flood_no.txt")

```

J.4B Descriptive Statistics
```{r}
# Calculate the number of observations for each group
n_obs_flood_yes <- nrow(flood_yes)
n_obs_flood_no <- nrow(flood_no)

# Make vectors of the mean and standard deviation for each variable in each group
mean_df_r <- colMeans(flood_yes, na.rm = TRUE)
sd_df_r <- apply(flood_yes, 2, sd, na.rm = TRUE)
mean_df_u <- colMeans(flood_no, na.rm = TRUE)
sd_df_u <- apply(flood_no, 2, sd, na.rm = TRUE)

# Make a vector of the number of observations for each variable (all the same for each group)
n_obs_flood_yes_vec <- rep(n_obs_flood_yes, ncol(flood_yes))
n_obs_flood_no_vec <- rep(n_obs_flood_no, ncol(flood_no))

# Calculate p-values for each variable using a two-sample t-test
pvalues <- numeric(ncol(flood_yes))

for (i in 1:ncol(flood_yes)) {
  t_test <- t.test(flood_yes[, i], flood_no[, i])
  pvalues[i] <- t_test$p.value
}

# Combine the means, standard deviations, p-values, and number of observations into a summary table
summary_table <- data.frame(
  Variable = colnames(flood_yes),
  Mean_Flood_Yes = mean_df_r,
  SD_Flood_Yes = sd_df_r,
  Mean_Flood_No = mean_df_u,
  SD_Flood_No = sd_df_u,
  N_Obs_Flood_Yes = n_obs_flood_yes_vec,
  N_Obs_Flood_No = n_obs_flood_no_vec,
  P_Value = pvalues
)

# Print the summary table to the console
print(summary_table)

# Save the summary table to a CSV file
write.csv(summary_table, "summary_statistics.csv")

```

J.5A Regression Analysis
```{r}
library(stargazer)
library(glm2)

# Covariates
base_formula <- " ~ flood_kde_1 + Duration + PermanentWater"
interaction_formula <- paste(base_formula, " + flood_kde_1:Duration + flood_kde_1:PermanentWater")
full_formula <- paste(interaction_formula, " + age + rural + female + wealth + factor(region)")

# List of dependent variables and their respective file outputs
dep_vars <- c("deprived_sev", "dep_education_sev", "dep_health_sev", "dep_housing_sev", "dep_nutrition_sev", "dep_sanitation_sev", "dep_water_sev")
file_names <- c("logit_floods_poor", "logit_floods_education", "logit_floods_health", "logit_floods_housing", "logit_floods_nutrition", "logit_floods_sanitation", "logit_floods_water")

# Loop through each dependent variable
for (i in 1:length(dep_vars)) {
  
  # Base regression
  logit_1 <- glm(paste(dep_vars[i], base_formula), data = dhs_flood, family = binomial)
  
  # Interaction regression
  logit_2 <- glm(paste(dep_vars[i], interaction_formula), data = dhs_flood, family = binomial)
  
  # Full regression with region fixed effects
  logit_3 <- glm(paste(dep_vars[i], full_formula), data = dhs_flood, family = binomial)
  
  # Generate the table
  stargazer(logit_1, logit_2, logit_3, 
            title=paste("Logistic Regression -", dep_vars[i]), 
            align=TRUE,
            covariate.labels=c("Flood KDE", "Flood Duration", "Permanent Water", "Age", "Rural", "Female", "Wealth", "Flood KDE x Duration", "Flood KDE x PermanentWater"),
            dep.var.labels=dep_vars[i],
            out=paste(file_names[i], ".txt", sep=""),
            apply.coef=exp,
            omit=c("Constant", "factor(region)*"),  # omit regional FE coefficients and the constant
            notes=c("Coefficients are exponentiated", "Third regression includes regional FE"))
}

```
J.5A Regression Analysis + Survey Weights
```{r}
library(stargazer)
library(glm2)

# Covariates
base_formula <- " ~ flood_kde_1 + Duration + PermanentWater"
interaction_formula <- paste(base_formula, " + flood_kde_1:Duration + flood_kde_1:PermanentWater")
full_formula <- paste(interaction_formula, " + age + rural + female + wealth + factor(region)")

# List of dependent variables, their respective file outputs, and descriptive labels
dep_vars <- c("deprived_sev", "dep_education_sev", "dep_health_sev", "dep_housing_sev", "dep_nutrition_sev", "dep_sanitation_sev", "dep_water_sev")
file_names <- c("logit_floods_poor", "logit_floods_education", "logit_floods_health", "logit_floods_housing", "logit_floods_nutrition", "logit_floods_sanitation", "logit_floods_water")
dep_var_labels <- c("Severe Deprivation", "Education Deprivation", "Health Deprivation", "Housing Deprivation", "Nutrition Deprivation", "Sanitation Deprivation", "Water Deprivation")

# Loop through each dependent variable
for (i in 1:length(dep_vars)) {
  
  # Base regression
  logit_1 <- glm(paste(dep_vars[i], base_formula), data = dhs_flood, family = binomial)
  
  # Interaction regression
  logit_2 <- glm(paste(dep_vars[i], interaction_formula), data = dhs_flood, family = binomial)
  
  # Full regression with region fixed effects
  logit_3 <- glm(paste(dep_vars[i], full_formula), data = dhs_flood, family = binomial)
  
  # Generate the table
  stargazer(logit_1, logit_2, logit_3, 
            title=paste("Logistic Regression -", dep_var_labels[i]), 
            align=TRUE,
            covariate.labels=c("Flood KDE", "Flood Duration", "Permanent Water", "Age", "Rural", "Female", "Wealth", "Flood KDE x Duration", "Flood KDE x PermanentWater"),  
            dep.var.labels=dep_var_labels[i], 
            out=paste(file_names[i], ".txt", sep=""),
            apply.coef=exp,
            omit=c("Constant", "region"),  # Excluding region from the output
            notes="Coefficients are exponentiated")
}

```

J.6 Propensity Score Matching
```{r}
library(stargazer)
library(broom)
library(Matching)
library(dplyr)
library(tidyverse)
library(broom)

# Create the new variable "rural"
dhs_flood$rural <- ifelse(dhs_flood$URBAN_RURA == "R", 1, 0)

# Create the new variable "female"
dhs_flood$female <- ifelse(dhs_flood$sex == "Female", 1, 0)

# Create a subset for only household located in flooded areas
dhs_flood <- subset(dhs_flood, flood_kde_11 != 0)

# flooding variable
dhs_flood$flood <- ifelse(dhs_flood$flood_kde_1!=0,1,0)

# Logistic regression to estimate propensity scores
propensity_model <- glm(flood ~ age + rural + female + wealth, 
                        data = dhs_flood, family = binomial)

# Extract propensity scores
dhs_flood$propensity_score <- predict(propensity_model, type = "response")

# Perform matching based on propensity scores
matched_data <- Match(
  Y = dhs_flood$deprived_sev,
  Tr = dhs_flood$flood,
  X = dhs_flood$propensity_score
)

# Extract the matched dataset
matched_dhs_flood <- dhs_flood[unlist(matched_data[c("index.treated", "index.control")]), ]

# List of outcomes
outcomes <- c("deprived_sev", 
              "dep_education_sev", 
              "dep_health_sev", 
              "dep_housing_sev",
              "dep_nutrition_sev", 
              "dep_sanitation_sev", 
              "dep_water_sev")

# Create a named list that maps each outcome variable to a descriptive label
outcome_labels <- c(
  deprived_sev = "Severely Deprived",
  dep_education_sev = "Deprived in Education",
  dep_health_sev = "Deprived in Health",
  dep_housing_sev = "Deprived in Housing",
  dep_nutrition_sev = "Deprived in Nutrition",
  dep_sanitation_sev = "Deprived in Sanitation",
  dep_water_sev = "Deprived in Water"
)

# Loop through outcomes and perform logistic regression
for (outcome in outcomes) {
  
  # Define the formula for the logistic regression
  formula_m1 <- as.formula(paste(outcome, "~ flood"))
  formula_m2 <- as.formula(paste(outcome, "~ flood + age + rural + female + wealth"))
  
  # Perform the logistic regression for each model
  logit_m1_matched <- glm(formula_m1, data = matched_dhs_flood, family = binomial)
  logit_m2_matched <- glm(formula_m2, data = matched_dhs_flood, family = binomial)
  
  # Fetch the appropriate label for the current outcome
  current_label <- outcome_labels[outcome]
  
  # Generate the table for the current outcome (matched data)
  stargazer(logit_m1_matched, logit_m2_matched, 
            title=paste("Logistic Regression -", outcome, "(Matched Data)"), 
            align=TRUE,
            covariate.labels=c("Flood KDE", "Age", "Rural", "Female", "Wealth"),
            dep.var.labels=current_label, # Use the fetched label
            out=paste("logit_floods_", outcome, "_matched.txt", sep=""),
            apply.coef=exp,
            omit="Constant",
            notes="Coefficients are exponentiated")
}

```

J.7 Balance Check - PSM
```{r}
library(writexl)

# Function to calculate mean, SD, N, and p-value for SMD
get_balance_statistics <- function(data, treatment_var, covariates) {
  treated <- data[data[[treatment_var]] == 1, ]
  control <- data[data[[treatment_var]] == 0, ]
  
  means_treated <- sapply(covariates, function(var) mean(treated[[var]], na.rm = TRUE))
  means_control <- sapply(covariates, function(var) mean(control[[var]], na.rm = TRUE))
  
  sd_treated <- sapply(covariates, function(var) sd(treated[[var]], na.rm = TRUE))
  sd_control <- sapply(covariates, function(var) sd(control[[var]], na.rm = TRUE))
  
  n_treated <- length(treated[[treatment_var]])
  n_control <- length(control[[treatment_var]])
  
  sd_pooled <- sapply(covariates, function(var) sqrt(((n_treated - 1) * sd_treated[var] + 
                                                      (n_control - 1) * sd_control[var]) / 
                                                     (n_treated + n_control - 2)))
  
  SMD <- (means_treated - means_control) / sd_pooled
  
  # Calculate the p-value for SMD using Z-test
  Z <- SMD / sqrt(2/n_treated) # Assuming equal sample size for treated and control for simplicity
  p_value <- 2 * (1 - pnorm(abs(Z)))
  
  return(data.frame(Covariate = covariates,
                    Mean_Treated = means_treated, SD_Treated = sd_treated, N_Treated = n_treated,
                    Mean_Control = means_control, SD_Control = sd_control, N_Control = n_control,
                    SMD = SMD, P_Value = p_value))
}

# Calculate balance statistics for unmatched data
balance_unmatched <- get_balance_statistics(dhs_flood, "flood", covariates)

# Calculate balance statistics for matched data
balance_matched <- get_balance_statistics(matched_dhs_flood, "flood", covariates)

# Combine both tables for easy comparison
combined_balance <- rbind(cbind(balance_unmatched, Dataset = "Unmatched"),
                         cbind(balance_matched, Dataset = "Matched"))

combined_balance

write_xlsx(combined_balance, "balance_statistics.xlsx")
```

