---
Title: "Studying the relationship betwen climate change and children's wellbeing"
Author: Hernando Grueso
Output: climate_children
---

Creation: 09/05/2023
Update: 08/08/2023
Data Sources: 
1) Flood maps from Nigeria: https://global-flood-database.cloudtostreet.ai/
2) DHS:
3) VACS:
4) MICS:

A. Install Required Packages
```{r}
# Install Required Packages
if(!requireNamespace("raster")) install.packages( "raster" )
if(!requireNamespace("MASS")) install.packages( "MASS" )
if(!requireNamespace("rgdal")) install.packages( "rgdal" )
if(!requireNamespace("spatstat")) install.packages( "spatstat" )
if(!requireNamespace("sp")) install.packages( "sp" )
if(!requireNamespace("rnaturalearth")) install.packages( "rnaturalearth" )
if(!requireNamespace("rnaturalearthdata")) install.packages( "rnaturalearthdata" )
if(!requireNamespace("dplyr")) install.packages( "dplyr" )
if(!requireNamespace("sf")) install.packages( "sf" )
if(!requireNamespace("RColorBrewer")) install.packages( "RColorBrewer" )
if(!requireNamespace("leaflet")) install.packages( "leaflet" )
if(!requireNamespace("htmlwidgets")) install.packages( "htmlwidgets" )
if(!requireNamespace("ggplot2")) install.packages( "ggplot2" )
if(!requireNamespace("stargazer")) install.packages( "stargazer" )
```

B. Define Directory
```{r}
# Defining Directory:
groupDir <- "C:/Users/hgruesohurtado/Desktop/Climate Change"

```

C. Import Data
```{r}
# Import GeoTiff file
library(raster)
raster_layer <- raster("Floods/2018/DFO_4667_From_20180820_to_20181002.tif") 
#raster_layer_16 <- raster("Floods/2016/DFO_4391_From_20160914_to_20160922.tif") 
#raster_layer <- raster("Floods/2015/DFO_4294_From_20150830_to_20150911.tif") 

# Convert to a dataframe
df <- as.data.frame(raster_layer, xy=TRUE)

# Import DHS geo-codes
library(readr)
dhs <- read_csv("Save the Children/Child poverty and climate/dhsgps_combined_sep22.csv", show_col_types = FALSE)
dhs <- subset(dhs, dhs$DHSYEAR==2018)

# Import DHS data 2018
dhs_df <- read_csv("Save the Children/Child poverty and climate/Microdata_DHS_Apr2023.csv", show_col_types = FALSE)
dhs_df <- subset(dhs_df, dhs_df$year_interview==2018)
```

D. Merge DHS MPI & GEO-Codes
```{r}
# Only keep Nigeria in DHS
dhs <- subset(dhs, dhs$DHSCC=="NG")
dhs_df <- subset(dhs_df, dhs_df$countrycode=="NGA")

# Merge at the cluster level (geo-codes are defined at the cluster level)
dhs_df <- merge(dhs_df, dhs, by = "cluster")

# Delete data frames that will not be used
rm(df)
```

E. Plot Nigeria flooding map
```{r}
# Plot the raster map with custom settings
plot(raster_layer,
     main = "Foold Map: Nigeria, 2018",
     legend = TRUE,
     legend.args = list(text = "Flooding"))
```

F.1 Plot Severe MPI=1 DHS households on flooding map  
```{r}
# Plot the raster layer
plot(raster_layer,
     main = "Flood Map: Nigeria, 2018 (Child Deprived: Severe)",
     legend = TRUE,
     legend.args = list(text = "Flooding"))

# Create subset of severe poor households
dhs_poor_sev <- subset(dhs_df, dhs_df$deprived_sev==1)
  
# Add DHS data points to the plot
points(dhs_poor_sev$LONGNUM, dhs_poor_sev$LATNUM, col = "blue", pch = 20)
```

F.2 Plot Moderate MPI=0 DHS households on flooding map  
```{r}
# Plot the raster layer
plot(raster_layer,
     main = "Flood Map: Nigeria, 2018 (Child Non-Deprived: Moderate)",
     legend = TRUE,
     legend.args = list(text = "Flooding"))

# Create subset of moderately poor households
dhs_poor_mod <- subset(dhs_df, dhs_df$deprived_mod==0)
  
# Add DHS data points to the plot
points(dhs_poor_mod$LONGNUM, dhs_poor_mod$LATNUM, col = "purple", pch = 20)

```

G.1 Plot Severe MPI=0 DHS households on flooding map  
```{r}
# Plot the raster layer
plot(raster_layer,
     main = "Flood Map: Nigeria, 2018 (Child Non-Deprived: Severe)",
     legend = TRUE,
     legend.args = list(text = "Flooding"))

# Create subset of severe poor households
dhs_poor_sev <- subset(dhs_df, dhs_df$deprived_sev==0)
  
# Add DHS data points to the plot
points(dhs_poor_sev$LONGNUM, dhs_poor_sev$LATNUM, col = "blue", pch = 20)

```

G.2 Plot Flooding Map & DHS Households Deprived on: Housing (moderate)  
```{r}
# Plot the raster layer
plot(raster_layer,
     main = "Flood Map: Nigeria, 2018 (Housing Deprivation: Moderate)",
     legend = TRUE,
     legend.args = list(text = "Flooding"))

# Create subset of moderately poor households
dhs_poor_mod <- subset(dhs_df, dhs_df$dep_housing_mod==1)
  
# Add DHS data points to the plot
points(dhs_poor_mod$LONGNUM, dhs_poor_mod$LATNUM, col = "orange", pch = 20)

```

H.1 Plot Flooding Map & DHS Households Deprived on: Water
```{r}
# Plot the raster layer
plot(raster_layer,
     main = "Flood Map: Nigeria, 2018 (Water Deprivation: Moderate)",
     legend = TRUE,
     legend.args = list(text = "Flooding"))

# Create subset of moderately poor households
dhs_poor_mod <- subset(dhs_df, dhs_df$dep_water_mod==1)
  
# Add DHS data points to the plot
points(dhs_poor_mod$LONGNUM, dhs_poor_mod$LATNUM, col = "blue", pch = 20)

```

H.2 Plot Flooding Map & DHS Households Deprived on: Sanitation
```{r}
# Plot the raster layer
plot(raster_layer,
     main = "Flood Map: Nigeria, 2018 (Sanitation Deprivation: Moderate)",
     legend = TRUE,
     legend.args = list(text = "Flooding"))

# Create subset of moderately poor households
dhs_poor_mod <- subset(dhs_df, dhs_df$dep_sanitation_mod==1)
  
# Add DHS data points to the plot
points(dhs_poor_mod$LONGNUM, dhs_poor_mod$LATNUM, col = "red", pch = 20)

```

H.3 Plot Flooding Map & DHS Households Deprived on: Nutrition
```{r}
# Plot the raster layer
plot(raster_layer,
     main = "Flood Map: Nigeria, 2018 (Nutrition Deprivation: Moderate)",
     legend = TRUE,
     legend.args = list(text = "Flooding"))

# Create subset of moderately poor households
dhs_poor_mod <- subset(dhs_df, dhs_df$dep_nutrition_mod==1)
  
# Add DHS data points to the plot
points(dhs_poor_mod$LONGNUM, dhs_poor_mod$LATNUM, col = "purple", pch = 20)

```

H.4 Plot Flooding Map & DHS Households Deprived on: Health
```{r}
# Plot the raster layer
plot(raster_layer,
     main = "Flood Map: Nigeria, 2018 (Health Deprivation: Moderate)",
     legend = TRUE,
     legend.args = list(text = "Flooding"))

# Create subset of moderately poor households
dhs_poor_mod <- subset(dhs_df, dhs_df$dep_health_mod==1)
  
# Add DHS data points to the plot
points(dhs_poor_mod$LONGNUM, dhs_poor_mod$LATNUM, col = "orange", pch = 20)

```

H.6 Plot Flooding Map & DHS Households Deprived on: Education
```{r}
# Plot the raster layer
plot(raster_layer,
     main = "Flood Map: Nigeria, 2018 (Education Deprivation: Moderate)",
     legend = TRUE,
     legend.args = list(text = "Flooding"))

# Create subset of moderately poor households
dhs_poor_mod <- subset(dhs_df, dhs_df$dep_education_mod==1)
  
# Add DHS data points to the plot
points(dhs_poor_mod$LONGNUM, dhs_poor_mod$LATNUM, col = "yellow", pch = 20)

```

I.1 Map of Poverty Severity at the Cluster Level
```{r}
# Load necessary packages
library(raster)
library(ggplot2)
library(scales)
library(RColorBrewer)

# Create a color blending function
color_blend_func <- colorRampPalette(colors = c("green", brewer.pal(9, "RdYlBu")))

# Plot the raster data with a lighter color palette
plot(raster_layer, col = color_blend_func(256), axes = TRUE, main = 'Flood Map: Nigeria, 2018 with Cluster Poverty Severity')

# Create a color palette
color_func <- colorRampPalette(colors = c("red", "red"))

# Calculate color indices
color_indices <- round(dhs_df$deprived_sev * 100)

# Ensure color indices are within valid range
color_indices[color_indices > 100] <- 100
color_indices[color_indices < 1] <- 1

# Generate colors from palette
colors <- alpha(color_func(100)[color_indices], 0.7)

# Add the clusters' locations to the plot with small size and red color
points(dhs_df$LONGNUM, dhs_df$LATNUM, col = colors, pch = 20, cex = 0.5)
```

I.2 Geo-Spatial Distribution of Poverty
```{r}
library(tidyverse)

# Collapse poverty insidence at the cluster level
dhs_cluster <- dhs_df %>%
  filter(LATNUM != 0 & LONGNUM != 0) %>%
  group_by(cluster) %>%
  summarize(avg_deprived_sev = mean(deprived_sev),
            LATNUM = first(LATNUM),
            LONGNUM = first(LONGNUM))

# Plot the data
library(ggplot2)
ggplot(dhs_cluster, aes(x = LONGNUM, y = LATNUM)) +
  geom_point(aes(color = avg_deprived_sev)) +
  scale_color_gradient(low = "yellow", high = "red") +
  labs(title = "Geo-Spatial Distribution of Poverty",
       x = "Longitude", 
       y = "Latitude",
       color = "Average Deprivation Level") +
  theme_minimal()
```
I.3 Geo-Spatial Distribution of Wealth
```{r}
# Collapse poverty incidence at the cluster level
dhs_cluster <- dhs_df %>%
  filter(LATNUM != 0 & LONGNUM != 0) %>%
  group_by(cluster) %>%
  summarize(wealth = mean(wealth),
            LATNUM = first(LATNUM),
            LONGNUM = first(LONGNUM))

# Load the necessary library
library(ggplot2)

# Check the data
head(dhs_cluster)

# Plot the data
ggplot(dhs_cluster, aes(x = LONGNUM, y = LATNUM)) +
  geom_point(aes(color = wealth)) +
  scale_color_gradient(low = "red", high = "yellow") +
  labs(title = "Geo-Spatial Distribution of Wealth",
       x = "Longitude", 
       y = "Latitude",
       color = "Average Wealth Level") +
  theme_minimal()
```

J.1 Kernel Density Estimate
```{r}
# Convert the raster to points
points <- rasterToPoints(raster_layer)

# Perform kernel density estimation
kde <- density(points)

# Plot the kernel density estimate
plot(kde)

```

J.2 Calculate KDE 1x1 and 11x11
```{r}
library(raster)
library(RColorBrewer)
library(sp)
library(lattice)
library(grid)

# Convert the raster to a binary mask
flooded_mask <- raster_layer > 0

# Calculate density with 1x1 bandwidth
bandwidth1 <- matrix(1, nrow = 1, ncol = 1)
density1 <- focal(flooded_mask, w = bandwidth1, fun = sum)

# Calculate density with 41x41 bandwidth
bandwidth11 <- matrix(1, nrow = 11, ncol = 11)
density11 <- focal(flooded_mask, w = bandwidth11, fun = sum)

# Identify the raster with fewer non-NA values
if (sum(!is.na(density1[])) < sum(!is.na(density11[]))) {
  mask_raster <- density1
} else {
  mask_raster <- density11
}

# Use the mask_raster to mask both density rasters
density1_aligned <- mask(density1, mask_raster)
density11_aligned <- mask(density11, mask_raster)

# Combine the two density rasters into a multilayer raster
density_stack <- stack(density1_aligned, density11_aligned)

# Create a custom multi-color palette
my_palette <- colorRampPalette(c(brewer.pal(3, "Blues"), brewer.pal(3, "Greens"), brewer.pal(3, "Reds")))

# Create a custom plot function using levelplot (from the lattice package)
# Plot the two density rasters with the custom color palette
spplot_obj <- spplot(density_stack, names.attr = c("Bandwidth 1x1", "Bandwidth 11x11"), col.regions = my_palette(100))

# Convert the rasters to points (data frames)
density1_points <- rasterToPoints(density1_aligned)
density11_points <- rasterToPoints(density11_aligned)

# Combine into one data frame
df_flood18 <- data.frame(
  Longitude = density1_points[,1],
  Latitude = density1_points[,2],
  Density1 = density1_points[,3],
  Density11 = density11_points[,3]
)

# Print the spplot object
print(spplot_obj)

# Add a title
grid.text("Kernel Density Estimates: Flooding Nigeria, 2018", x = 0.5, y = 0.98, gp = gpar(fontsize = 15))
```

J.3 Merge KDE and DHS
```{r}
# Rename variable before merging
names(df_flood18)[names(df_flood18) == "Longitude"] <- "LONGNUM"
names(df_flood18)[names(df_flood18) == "Latitude"] <- "LATNUM"
names(df_flood18)[names(df_flood18) == "Density1"] <- "flood_kde_1"
names(df_flood18)[names(df_flood18) == "Density11"] <- "flood_kde_11"

# Merge at the cluster level (geo-codes are defined at the cluster level)
dhs_flood <- merge(dhs_df, df_flood18, by = c("LATNUM", "LONGNUM"))

# Convert datasets to sf objects
library(sf)
dataset1_sf <- st_as_sf(dhs_df, coords = c("LONGNUM", "LATNUM"))
dataset2_sf <- st_as_sf(df_flood18, coords = c("LONGNUM", "LATNUM"))

# Perform a spatial join based on proximity, only include matched observations
dhs_flood <- st_join(dataset1_sf, dataset2_sf, join = st_nearest_feature, left = FALSE, dist = 100)  # Adjust the distance threshold as needed

# Check the number of approximate matches
num_approx_matches <- nrow(dhs_flood)
print(dhs_flood)
```

J.4 Descriptive Statistics
```{r}
library(stargazer)
library(dplyr)

# Create the new variable "rural"
dhs_flood$rural <- ifelse(dhs_flood$URBAN_RURA == "R", 1, 0)

# Create the new variable "female"
dhs_flood$female <- ifelse(dhs_flood$sex == "Female", 1, 0)

# Create a subset for only hosehold located in flooded areas
dhs_flood <- subset(dhs_flood, flood_kde_11 != 0)

# Subset
flood_yes <- dhs_flood[dhs_flood$flood_kde_1==1, ]
flood_no <- dhs_flood[dhs_flood$flood_kde_1==0, ]

flood_yes <- flood_yes[c("deprived_sev", "deprived_mod", "age", "rural", "female", "wealth")]
flood_no <- flood_no[c("deprived_sev", "deprived_mod", "age", "rural", "female", "wealth")]

flood_yes$geometry <- NULL
flood_no$geometry <- NULL

# Compute descriptive stats
mean_df_r <- sapply(flood_yes, mean)
sd_df_r <- sapply(flood_yes, sd)
obs_df_r <- sapply(flood_yes, function(x) sum(!is.na(x)))

mean_df_u <- sapply(flood_no, mean)
sd_df_u <- sapply(flood_no, sd)
obs_df_u <- sapply(flood_no, function(x) sum(!is.na(x)))

stargazer(as.data.frame(flood_yes), 
          type = "text", 
          title="Descriptive Stats", 
          digits=2, 
          out="desc_stats_flood_yes.txt")

stargazer(as.data.frame(flood_no), 
          type = "text", 
          title="Descriptive Stats", 
          digits=2, 
          out="desc_stats_flood_no.txt")

```

J.5 Regression Analysis
```{r}
# Load the packages
library(stargazer)
library(broom)

# Perform the logistic regression analysis with fixed effects and covariates
logit_e1 <- glm(dep_education_mod ~ flood_kde_1, data = dhs_flood, family = binomial)
logit_e2 <- glm(dep_education_mod ~ flood_kde_1 + age + rural + female + wealth, data = dhs_flood, family = binomial)

logit_he1 <- glm(dep_health_mod ~ flood_kde_1, data = dhs_flood, family = binomial)
logit_he2 <- glm(dep_health_mod ~ flood_kde_1 + age + rural + female + wealth, data = dhs_flood, family = binomial)

logit_ho1 <- glm(dep_housing_mod ~ flood_kde_1, data = dhs_flood, family = binomial)
logit_ho2 <- glm(dep_housing_mod ~ flood_kde_1 + age + rural + female + wealth, data = dhs_flood, family = binomial)

logit_n1 <- glm(dep_nutrition_mod ~ flood_kde_1, data = dhs_flood, family = binomial)
logit_n2 <- glm(dep_nutrition_mod ~ flood_kde_1 + age + rural + female + wealth, data = dhs_flood, family = binomial)

logit_s1 <- glm(dep_sanitation_mod ~ flood_kde_1, data = dhs_flood, family = binomial)
logit_s2 <- glm(dep_sanitation_mod ~ flood_kde_1 + age + rural + female + wealth, data = dhs_flood, family = binomial)

logit_w1 <- glm(dep_water_mod ~ flood_kde_1, data = dhs_flood, family = binomial)
logit_w2 <- glm(dep_water_mod ~ flood_kde_1 + age + rural + female + wealth, data = dhs_flood, family = binomial)

# Generate the table for the education deprivation models
stargazer(logit_e1, logit_e2, 
          title="Logistic Regression Results - Education Deprivation", 
          align=TRUE,
          covariate.labels=c("Flood KDE", "Age", "Rural", "Female", "Wealth"),
          dep.var.labels.include=FALSE,
          out="logit_floods_education.txt",
          apply.coef = exp)

# Generate the table for the health deprivation models
stargazer(logit_he1, logit_he2,
          title="Logistic Regression Results - Health Deprivation", 
          align=TRUE,
          covariate.labels=c("Flood KDE", "Age", "Rural", "Female", "Wealth"),
          dep.var.labels.include=FALSE,
          out="logit_floods_health.txt",
          apply.coef = exp)

# Generate the table for the housing deprivation models
stargazer(logit_ho1, logit_ho2, 
          title="Logistic Regression Results - Housing Deprivation", 
          align=TRUE,
          covariate.labels=c("Flood KDE", "Age", "Rural", "Female", "Wealth"),
          dep.var.labels.include=FALSE,
          out="logit_floods_housing.txt",
          apply.coef = exp)

# Generate the table for the nutrition deprivation models
stargazer(logit_n1, logit_n2, 
          title="Logistic Regression Results - Nutrition Deprivation", 
          align=TRUE,
          covariate.labels=c("Flood KDE", "Age", "Rural", "Female", "Wealth"),
          dep.var.labels.include=FALSE,
          out="logit_floods_nutrition.txt",
          apply.coef = exp)

# Generate the table for the sanitation deprivation models
stargazer(logit_s1, logit_s2, 
          title="Logistic Regression Results - Sanitation Deprivation", 
          align=TRUE,
          covariate.labels=c("Flood KDE", "Age", "Rural", "Female", "Wealth"),
          dep.var.labels.include=FALSE,
          out="logit_floods_sanitation.txt",
          apply.coef = exp)

# Generate the table for the water deprivation models
stargazer(logit_w1, logit_w2, 
          title="Logistic Regression Results - Water Deprivation", 
          align=TRUE,
          covariate.labels=c("Flood KDE", "Age", "Rural", "Female", "Wealth"),
          dep.var.labels.include=FALSE,
          out="logit_floods_water.txt",
          apply.coef = exp)
```
